Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-13 qhtv2.tex
wip

\subsection{Bounds on the Sliding Window DDP}

@@

However, this implementation depends on $\log_2 |\Gamma|$, which might be arbitrarily large. We thus propose another solution, with a very low error rate but constant time.

@@

	More precisely, it is possible to create a filter of $M$ bits with an FNR of $0$, and an FPR of $1 - (1-\frac1{w^2})^w \sim \frac 1w$.
\end{theorem}

@@

Let $\mathcal F$ be the following DDF: the filter's state consists in a queue of $w$ hashes, and for each new element $e$, $\mathsf{Detect}(e)$ returns \DUPLICATE if $h(e)$ is present in the queue, \UNSEEN otherwise. $\mathsf{Insert}(e)$ appends $h(e)$ to the queue before popping the list.

There is no false negative, and a false positive only happens if the new element to be inserted collides with at least one other element, which happens with probability $1 - (1 - \frac 1{2^{2\log_2w}})^{w} = 1 - (1-\frac1{w^2})^w$, hence an FNR of $0$ and a FPR of $1 - (1-\frac1{w^2})^w$.
The queue stores $w$ hashes, and as such requires $w \cdot 2\log_2 w$ bits of memory.
\end{proof}

@@

Note that this solution has a time complexity of $O(w)$. Using a dictionary like in the previous proof, we get a filter with an error rate of about $\frac 1w$ and constant time for insertion and lookup, but requiring $5w\log_2 w$ bits of memory.

@@

\subsection{DDP vs Sliding Window DDP}
The sliding window variant of DDP may seem to be a slight weakening of the original problem. Remarkably, it turns out that DDFs which perform well for the DDP do rather poorly for the sliding window DDP. A literature review collects the following DDF constructions: A2 filters \cite{Yoo10}, Stable Bloom Filters (SBF) \cite{Den06}, Quotient Hash Tables (QHT) \cite{10.1145/3297280.3297335}, Streaming Quotient Filters (SQF)\footnote{QHTs stem from a correction of SQFs, and as such SQFs are less efficient than QHT in every situation \cite{10.1145/3297280.3297335}. We included them for the sake of completeness.} \cite{Dut13}, Block-decaying Bloom Filters (b\_DBF) \cite{She08}, and a slight variation of Cuckoo Filters \cite{Fan14} suggested by \cite{10.1145/3297280.3297335}. The structure proposed in \cite{10.1145/1060745.1060753} is not suited to either sliding window or no sliding window DDP, as they work on what they call `landmark` sliding window, which is a zero-resetting of the memory at some user-defined epochs.

@@

However, practical tests show (see \ref{sub:L_bench}) that setting $L = \sqrt w$ leads to optimal results.

Hence, the (heuristic) optimal solution is to set $L = \sqrt w$.

@@

\subsection{Saturation Resistance of DDFs}\label{sub:saturation}
In this section, we compare the resistance of various filters to saturation. We take the same filters and parameters as \cite{10.1145/3297280.3297335}, namely:

\begin{itemize}
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, 3 bits per fingerprint.
	\item SQF \cite{Dut13}, 1 bucket per row, $r = 2$ and $r' = 1$
	\item Cuckoo Filter \cite{Fan14}, cells containing 1 element of 3 bits each
	\item Stable Bloom Filter (SBF) \cite{Den06}, 2 bits per cell, 2 hash functions, targeted FPR of 0.02
	\item $A2$ Filter \cite{Yoo10}, targeted FPR of $0.1$ on the sliding window.
	\item Block-Decaying Bloom Filter (b\_DBF) \cite{She08}, sliding window of 6000 elements.
\end{itemize}

@@

\begin{figure}
	\input{er_n_artificial}
	\caption{Error rate (times 100) of DDFs of 1Mb, depending on the size of the stream, without any sliding window. Hatched area represents over-optimal (impossible) values}.\label{fig:graph_n}
\end{figure}

\subsection{Benchmarking Various Queuing Filters}
In this section, we experimentally prove the heuristic used in \ref{sub:optimal_ddf}, that the best DDF operating on a full stream leads to the best queueing filter operating on a sliding window.

>>>>

\subsection{Bounds on the Sliding Window DDP}\label{sub:opt_solutions}

@@

However, this implementation depends on $\log_2 |\Gamma|$, which might be arbitrarily large. We thus propose another solution, still in constant time but with a with a very low error rate, and not depending on $|\Gamma|$.

@@

	More precisely, it is possible to create a filter of $M$ bits with an FNR of $0$, and an FPR of $1 - (1-\frac1{w^2})^w \sim \frac 1w$, and a time complexity of $O(w)$.

	If $M \simeq 5w\log_2 w$, then it is possible to upgrade this filter so its error rates remain the same but the complexity drops to constant time.
\end{theorem}

@@

Let $\mathcal F$ be the following DDF: the filter's state consists in a queue of $w$ hashes, and for each new element $e$, $\mathsf{Detect}(e)$ returns \DUPLICATE if $h(e)$ is present in the queue, \UNSEEN otherwise. $\mathsf{Insert}(e)$ appends $h(e)$ to the queue before popping the queue.

There is no false negative, and a false positive only happens if the new element to be inserted collides with at least one other element, which happens with probability $1 - (1 - \frac 1{2^{2\log_2w}})^{w} = 1 - (1-\frac1{w^2})^w$, hence an FNR of $0$ and a FPR of $1 - (1-\frac1{w^2})^w$.
The queue stores $w$ hashes, and as such requires $w \cdot 2\log_2 w$ bits of memory.

Note that this solution has a time complexity of $O(w)$. Using an additional dictionary like in the previous proof, we get a filter with an error rate of about $\frac 1w$ and constant time for insertion and lookup, but requiring $5w\log_2 w$ bits of memory.
\end{proof}

@@

@@

\subsection{DDP vs Sliding Window DDP}
The sliding window variant of DDP may seem to be a slight weakening of the original problem. Remarkably, it turns out that DDFs which perform well for the DDP do rather poorly for the sliding window DDP. A literature review collects the following DDF constructions: A2 filters \cite{Yoo10}, Stable Bloom Filters (SBF) \cite{Den06}, Quotient Hash Tables (QHT) \cite{10.1145/3297280.3297335}, Streaming Quotient Filters (SQF) \cite{Dut13}, Block-decaying Bloom Filters (b\_DBF) \cite{She08}, and a slight variation of Cuckoo Filters \cite{Fan14} suggested by \cite{10.1145/3297280.3297335}. The structure proposed in \cite{10.1145/1060745.1060753} is not suited to either sliding window or no sliding window DDP, as they work on what they call `landmark` sliding window, which consists of a zero-resetting of the memory at some user-defined epochs.

@@

\todo[FAUX, peut etre log ? However, practical tests show (see \ref{sub:L_bench}) that setting $L = \sqrt w$ leads to optimal results.]

%Hence, the (heuristic) optimal solution is to set $L = \sqrt w$.

@@

We remind the reader that for small sliding windows compared to the memory size, optimal or nearly optimal solutions can be implemented, as discussed in \ref{sub:opt_solutions}. However, since we must evaluate the number of duplicates in the streams we use, we cannot benchmark our filter in real-life situations, as doing so would require way too much memory for keeping track of the duplicates.

\subsection{Saturation Resistance of DDFs}\label{sub:saturation}
In this section, we compare the resistance of various filters to saturation. From the streaming filters we found in the literature, we only considered the filters not designed for operating on a sliding window, and parameters as \cite{10.1145/3297280.3297335}, namely:

\begin{itemize}
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, 3 bits per fingerprint.
	\item SQF \cite{Dut13}, 1 bucket per row, $r = 2$ and $r' = 1$
	\item Cuckoo Filter \cite{Fan14}, cells containing 1 element of 3 bits each
	\item Stable Bloom Filter (SBF) \cite{Den06}, 2 bits per cell, 2 hash functions, targeted FPR of 0.02
%	\item $A2$ Filter \cite{Yoo10}, targeted FPR of $0.1$ on the sliding window.
%	\item Block-Decaying Bloom Filter (b\_DBF) \cite{She08}, sliding window of 6000 elements.
\end{itemize}

@@

\begin{figure}
	\input{er_n_artificial}
	\caption{Error rate (times 100) of DDFs of 1Mb, depending on the size of the stream, without any sliding window. Hatched area represents over-optimal (impossible) values.}\label{fig:graph_n}
\end{figure}

We observe that the best filters are, in order, QHT, SQF, Cuckoo and SBF. We also observe that QHT and SQF have error rates relatively close to the lower bound, hence suggesting that these filters are close to optimality.

\subsection{Benchmarking Various Queuing Filters}
Now that DDF are compared on a full stream, we observe how each of them behave when they act as a subfilter in our queueing scheme. The goal is to empirically prove the assumption made in \ref{sub:optimal_ddf}, stating that the best (error-wise) DDF operating on a full stream leads to the best (error-wise) queueing filter operating on a sliding window. From \ref{fig:graph_n}, we observe that the best DDF for DDP on a full stream are, in order, QHT, SQF, Cuckoo, SQF.

For this, we used four different queueing filters, with $L = 10$ subfilters each, and with the following subfilters: 
\begin{itemize}
	\item QHT subfilter \cite{10.1145/3297280.3297335}, 1 bucket per row, 3 bits per fingerprint. 
	\item Cuckoo Filter \cite{Fan14}, adapted for streaming as recommanded by \cite{10.1145/3297280.3297335}, cells containing 1 element of 3 bits each
	\item SQF \cite{Dut13}, 1 bucket per row, $r = 2$ and $r' = 1$
	\item SBF \cite{Den06}, 2 bits per cell, 2 hash functions, targeted FPR of 0.02
\end{itemize}

We used a random stream of $1,000,000$ elements uniformly selected from an alphabet of size $|\Gamma| = 2^{16}$, a memory size $M=100,000$ bits, a sliding window of size $w = 10,000$. The error rates (defined as the sum of $\FNR^w$ and $\FPR^w$) are displayed in \ref{tab:comparing_queueing}. We see that in all cases, the QHT subfilter performs better or almost as well than any other subfilter, hence validating the heuristic assumption.

\begin{table}[]
	\begin{tabular}{rccccc}
			   & $w=1000$ & $w=5000$ & $w=10^4$ & $w=10^5$ & $w=10^6$ \\ \hline
		QHT    & \textbf{9.84} 	  & \textbf{26.25}	 & \textbf{41.09}	 & \textbf{87.89}	  & 70.76      \\
		Cuckoo & 84.99	  & 99.94	 & 99.95	 & 99.93	  & 99.80      \\
		SQF    & 16.73 	  & 48.76	 & 69.25	 & 95.77 	  & \textbf{70.14}     \\
		SBF    & 54.88 	  & 93.93	 & 97.33	 & 97.41	  & 71.17      \\
	\end{tabular}
	\caption{Error rate (times 100) of various queueing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{16}$. For each sliding window, the best performance is written in bold.}\label{tab:comparing_queueing}
\end{table}
